---
title: "Chapter 3 - Bayesian SMS Spam Classifier"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

This example uses the [SMS spam collection](http://www.dt.fee.unicamp.br/~tiago/smsspamcollection/) created by Tiago A. Almeida and José María Gómez Hidalgo.

## Install and load packages

```
# Please run these by hand 
install.packages("tm")
install.packages("SnowballC")
install.packages("wordcloud") 
install.packages("e1071")
install.packages("gmodels")
```

```{r}
library(tm, quietly = TRUE)
library(SnowballC, quietly = TRUE)
library(wordcloud, quietly = TRUE)
library(e1071, quietly = TRUE)
library(gmodels, quietly = TRUE)
```

## Load the SMS data

```{r}
sms_raw <- read.csv("SMSSpamCollection.txt", stringsAsFactors = FALSE, sep = "\t", quote="", header = FALSE, col.names = c('type', 'text'))
sms_raw$type <- factor(sms_raw$type)  # convert type to a factor since there are only two types
```

## Create the Document Term Matrix

```{r}
sms_corpus <- VCorpus(VectorSource(sms_raw$text))
sms_dtm <- DocumentTermMatrix(sms_corpus, control = list(
    tolower = TRUE,
    removeNumbers = TRUE,
    stopwords = TRUE,
    removePunctuation = TRUE,
    stemming = TRUE
  ))

```

Split into training and test datasets

```{r}
sms_dtm_train <- sms_dtm[1:4169,]
sms_train_labels <- sms_raw[1:4169, ]$type

sms_dtm_test  <- sms_dtm[4170:5559,]
sms_test_labels  <- sms_raw[4170:5559, ]$type
```

## Compare wordclouds for spam vs ham

```{r}
spam <- subset(sms_raw, type == "spam")
ham <- subset(sms_raw, type == "ham")

wordcloud(spam$text, max.words = 40, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 40, scale = c(3, 0.5))
```

## Limit DTM to only have words that appear more than 5 times

```{r}
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
sms_dtm_freq_train<- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]

str(sms_freq_words)
```

## Convert DTM to categorical variables rather than word frequencies

```{r}
convert_counts <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)
```

## Train naive Bayes learner and check predictions on test dataset

```{r}
sms_classifier <- naiveBayes(sms_train, sms_train_labels, laplace=0)
sms_test_pred <- predict(sms_classifier, sms_test)
CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

## Train and check again using Laplace estimator

```{r}
sms_classifier2 <- naiveBayes(sms_train, sms_train_labels, laplace=1)
sms_test_pred2 <- predict(sms_classifier2, sms_test)
CrossTable(sms_test_pred2, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```